On Extending a Full-Sharing Design with Batched Scheduling

### 2015-04-15 paper.tex
batched scheduling ->ok for now

\begin{figure}[!ht]
\centering
\includegraphics[width=9.5cm]{figures/yap-mt.pdf}
\caption{Yap's multithreaded tabled space using the NS and FS designs}
\label{fig_yap_mt_support}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Batched Scheduling}

@@

Local scheduling strategy schedules the evaluation of a program in a
breath-first manner. It favors the backtracking first with completion
instead of the forward execution, leaving the consumption of answers
for last. Thus, it only allows a Cluster of Dependent Subgoals (CDS)
to return answers only after the completion point has been
reached~\cite{Freire-96}. In other words, the local scheduling tries
to keep a CDS as minimal as possible. When new answers are found, they
are added to the table space and the computation fails as consequence,
tabled subgoals inside a CDS propagate their answers to outside the
CDS only after its completion point is found. Local scheduling causes
a sooner completion of subgoals, which creates less complex
dependencies between them.

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=10.5cm]{figures/pcc.pdf}
\caption{The \emph{FS} design with the PAC procedure - (a) private
  chaining and (b) public chaining}
\label{fig_tabtries_pcc}
\end{figure}

@@

\begin{algorithm} [ht]
\caption{tabled\_new\_answer(answer ANS, subgoal frame SF)}
\algsetup{indent=0.65cm}
\begin{algorithmic}[1]
  \STATE $leaf \gets check\_insert\_answer\_trie(ANS, SF)$
  \STATE $chain \gets check\_insert\_consumer\_chain(leaf, SF)$
  \IF {$is\_answer\_marked\_as\_found(chain) = True$}
    \RETURN $failure$
  \ELSE                                        [the answer is new]
    \STATE $mark\_answer\_as\_found(chain)$
    \IF {$local\_scheduling\_mode(SF)$}
      \RETURN $failure$
    \ELSE                                [batched scheduling mode]
      \RETURN $proceed$
    \ENDIF
  \ENDIF  
\end{algorithmic}
\label{alg_table_new_answer_batched}
\end{algorithm}

@@

%% We used five sets of benchmarks. The \emph{Large Joins} and
%% \emph{WordNet} sets were obtained from the OpenRuleBench
%% project~\cite{Liang-09}; the \emph{Model Checking} set includes three
%% different specifications and transition relation graphs usually used
%% in model checking applications; the \emph{Path Left} and \emph{Path
%%   Right} sets implement two recursive definitions of the well-known
%% $path/2$ predicate, that computes the transitive closure in a graph,
%% using several different configurations of $edge/2$ facts. We
%% experimented graphs with the \emph{BTree} configuration with depth 17,
%% the \emph{Pyramid} and \emph{Cycle} configurations with depth 2000 and
%% the \emph{Grid} configuration with depth 35.

@@

\begin{table}[!ht]
\centering
\caption{Overhead ratios, when compared with the NS design with $1$
  thread (running with local scheduling without TabMalloc) for the NS
  and FS designs (with TabMalloc), when running 1, 8, 16, 24 and 32
  threads with local and batched scheduling on the five sets of
  benchmarks (best ratios by row and by design for the Minimum,
  Average and Maximum are in bold)}
%\scalebox{1.00} {
\begin{tabular}{ll|cc|cc}
\hline\hline
\multicolumn{2}{c|}{\multirow{2}{*}{\bf Threads}} &
\multicolumn{2}{c|}{\multirow{1}{*}{\bf NS}} &
\multicolumn{2}{|c}{\multirow{1}{*}{\bf FS}}\\ \cline{3-6}
& 
& \multicolumn{1}{c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}
& \multicolumn{1}{|c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}\\
\hline
\multirow{4}{*}{\bf 1}
& {\bf Min }& {\bf 0.53}& 0.55& 1.01& {\bf 0.95}\\
& {\bf Avg }& {\bf 0.78}& 0.82& {\bf 1.30}& 1.46\\
& {\bf Max }& 1.06& {\bf 1.05}& {\bf 1.76}& 2.33\\
& {\bf StD }& 0.15& 0.14& 0.22& 0.44\\
\hline
\multirow{4}{*}{\bf 8}
& {\bf Min }& 0.66& {\bf 0.63}& 1.16&{\bf  0.99}\\
& {\bf Avg }& {\bf 0.85}& 0.88& {\bf 1.88}& 1.95\\
& {\bf Max }& {\bf 1.12}& 1.14& {\bf 2.82}& 3.49\\
& {\bf StD }& 0.13& 0.14& 0.60& 0.79\\
\hline
\multirow{4}{*}{\bf 16}
& {\bf Min }& 0.85& {\bf 0.75}& 1.17& {\bf 1.06}\\
& {\bf Avg }& {\bf 0.98}& 1.00& {\bf 1.97}& 2.08\\
& {\bf Max }& {\bf 1.16}& 1.31& {\bf 3.14}& 3.69\\
& {\bf StD }& 0.09& 0.17& 0.65& 0.83\\
\hline
\multirow{4}{*}{\bf 24}
& {\bf Min }& {\bf 0.91}& 0.93& 1.16& {\bf 1.09}\\
& {\bf Avg }& {\bf 1.15}& 1.16& {\bf 2.06}& 2.19\\
& {\bf Max }& 1.72& {\bf 1.60}& {\bf 3.49}& 4.08\\
& {\bf StD }& 0.20& 0.21& 0.70& 0.91\\
\hline
\multirow{4}{*}{\bf 32}
& {\bf Min }& 1.05& {\bf 1.04}& 1.33& {\bf 1.26}\\
& {\bf Avg }& 1.51& {\bf 1.49}& {\bf 2.24}& 2.41\\
& {\bf Max }& {\bf 2.52}& 2.63& {\bf 3.71}& 4.51\\
& {\bf StD }& 0.45& 0.45& 0.74& 1.02\\
\hline\hline
\end{tabular}%}
\label{tab_batched_overhead}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Further Work}

>>>>

\begin{figure}[!ht]
\vspace{-\intextsep}
\centering
\includegraphics[width=9.5cm]{figures/yap-mt.pdf}
\caption{Yap's multithreaded tabled space using the NS and FS designs}
\label{fig_yap_mt_support}
\vspace{-1.0\intextsep}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Scheduling Strategies}

@@

Local scheduling strategy schedules the evaluation of a program in a
breath-first manner. It favors the backtracking first with completion
instead of the forward execution, leaving the consumption of answers
for last. Thus, it only allows a Cluster of Dependent Subgoals (CDS)
to return answers only after the completion point has been
reached~\cite{Freire-96}. %% In other words, the local scheduling tries
%% to keep a CDS as minimal as possible. When new answers are found, they
%% are added to the table space and the computation fails as consequence,
%% tabled subgoals inside a CDS propagate their answers to outside the
%% CDS only after its completion point is found.
Local scheduling causes a sooner completion of subgoals, which creates
less complex dependencies between them.

@@

\begin{figure}[!ht]
\vspace{-\intextsep}
\centering
\includegraphics[width=10.5cm]{figures/pcc.pdf}
\caption{The \emph{FS} design with the PAC procedure - (a) private
  chaining and (b) public chaining}
\label{fig_tabtries_pcc}
\vspace{-1.0\intextsep}
\end{figure}

@@

\begin{algorithm} [!ht]
\caption{tabled\_new\_answer(answer ANS, subgoal frame SF)}
\algsetup{indent=0.65cm}
\begin{algorithmic}[1]
  \STATE $leaf \gets check\_insert\_answer\_trie(ANS, SF)$
  \STATE $chain \gets check\_insert\_consumer\_chain(leaf, SF)$
  \IF {$is\_answer\_marked\_as\_found(chain) = True$}
    \RETURN $failure$
  \ELSE                                        [the answer is new]
    \STATE $mark\_answer\_as\_found(chain)$
    \IF {$local\_scheduling\_mode(SF)$}
      \RETURN $failure$
    \ELSE                                [batched scheduling mode]
      \RETURN $proceed$
    \ENDIF
  \ENDIF  
\end{algorithmic}
\label{alg_table_new_answer_batched}
\end{algorithm}

@@

We used five sets of benchmarks. The \emph{Large Joins} and
\emph{WordNet} sets were obtained from the OpenRuleBench
project~\cite{Liang-09}; the \emph{Model Checking} set includes three
different specifications and transition relation graphs usually used
in model checking applications; the \emph{Path Left} and \emph{Path
  Right} sets implement two recursive definitions of the well-known
$path/2$ predicate, that computes the transitive closure in a graph,
using several different configurations of $edge/2$ facts. We
experimented graphs with the \emph{BTree} configuration with depth 17,
the \emph{Pyramid} and \emph{Cycle} configurations with depth 2000 and
the \emph{Grid} configuration with depth 35.

@@

\begin{table}[!ht]
\vspace{-\intextsep}
\centering
\caption{Overhead ratios, when compared with the NS design with $1$
  thread (running with local scheduling without TabMalloc) for the NS
  and FS designs (with TabMalloc), when running 1, 8, 16, 24 and 32
  threads with local and batched scheduling on the five sets of
  benchmarks (best ratios by row and by design for the Minimum,
  Average and Maximum are in bold)}
\vspace{-0.5\intextsep}
%\scalebox{1.00} {
\begin{tabular}{ll|cc|cc}
\hline\hline
\multicolumn{2}{c|}{\multirow{2}{*}{\bf Threads}} &
\multicolumn{2}{c|}{\multirow{1}{*}{\bf NS}} &
\multicolumn{2}{|c}{\multirow{1}{*}{\bf FS}}\\ \cline{3-6}
& 
& \multicolumn{1}{c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}
& \multicolumn{1}{|c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}\\
\hline
\multirow{4}{*}{\bf 1}
& {\bf Min }& {\bf 0.53}& 0.55& 1.01& {\bf 0.95}\\
& {\bf Avg }& {\bf 0.78}& 0.82& {\bf 1.30}& 1.46\\
& {\bf Max }& 1.06& {\bf 1.05}& {\bf 1.76}& 2.33\\
& {\bf StD }& 0.15& 0.14& 0.22& 0.44\\
\hline
\multirow{4}{*}{\bf 8}
& {\bf Min }& 0.66& {\bf 0.63}& 1.16&{\bf  0.99}\\
& {\bf Avg }& {\bf 0.85}& 0.88& {\bf 1.88}& 1.95\\
& {\bf Max }& {\bf 1.12}& 1.14& {\bf 2.82}& 3.49\\
& {\bf StD }& 0.13& 0.14& 0.60& 0.79\\
\hline
\multirow{4}{*}{\bf 16}
& {\bf Min }& 0.85& {\bf 0.75}& 1.17& {\bf 1.06}\\
& {\bf Avg }& {\bf 0.98}& 1.00& {\bf 1.97}& 2.08\\
& {\bf Max }& {\bf 1.16}& 1.31& {\bf 3.14}& 3.69\\
& {\bf StD }& 0.09& 0.17& 0.65& 0.83\\
\hline
\multirow{4}{*}{\bf 24}
& {\bf Min }& {\bf 0.91}& 0.93& 1.16& {\bf 1.09}\\
& {\bf Avg }& {\bf 1.15}& 1.16& {\bf 2.06}& 2.19\\
& {\bf Max }& 1.72& {\bf 1.60}& {\bf 3.49}& 4.08\\
& {\bf StD }& 0.20& 0.21& 0.70& 0.91\\
\hline
\multirow{4}{*}{\bf 32}
& {\bf Min }& 1.05& {\bf 1.04}& 1.33& {\bf 1.26}\\
& {\bf Avg }& 1.51& {\bf 1.49}& {\bf 2.24}& 2.41\\
& {\bf Max }& {\bf 2.52}& 2.63& {\bf 3.71}& 4.51\\
& {\bf StD }& 0.45& 0.45& 0.74& 1.02\\
\hline\hline
\end{tabular}%}
\label{tab_batched_overhead}
\vspace{-0.7\intextsep}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Further Work}
.........
